Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-08-19 review_article.Rmd
captions for USA figures

```{r choro, fig.cap = "A choropleth map of the United States of America.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggchoro.png")
```

@@

!["Choropleth maps of the United States of America using four coordinate reference systems."](figures/ggchoroCRS.png)

@@

!["A contiguous cartogram of the Unites States of America"](figures/ggcont.png)

@@

!["A Non - contiguous cartogram of the Unites States of America"](figures/ggncont.png)

!["A dorling cartogram of the Unites States of America"](figures/ggdorl.png)

@@

He answers this with his implementation of maps created with 'the simplest of all shapes'. While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'. Circular cartograms use the same shape for every region represented, and size them according to the statistic represented or the population for a base map.
To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations. This implementation can work for up to 'one hundred thousand' areas.

<!--
-simple shapes of the circle, square, and triangle are relatively easily recognized.
-sacrifice communication efficiency;
-There is evidence, too, that angular shapes are more easily seen than ones with smooth outlines. Shapes that contain the qualities of simplicity and regularity are more readily perceived than others that do not possess these qualities.
-Finally, psychologists tell us that shapes having familiarity are more quickly
recognized than nonsense shapes or those that have no meaning.
-The cartographer can efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
-
@NISCC 
-->

Used in displays of the UK by Howe in 1986 cited by @HEDP
<!--
# Temporal displays
-->

>>>>

```{r choro, fig.cap = "A choropleth map of the United States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. The state of Utah has an unusually low rate of lung and bronchus incidence. The state of Kentucky is high, the surrounding states also have high rates. There is a pattern of average higher rates toward the East Coast and lower rates toward the West Coast.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggchoro.png")
```

@@

```{r choroCRS, fig.cap = "Four choropleth maps of the United States of America using various coordinate reference systems. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. The map projections alter the shapes and angles of the boundaries of each state. Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border. ", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggchoroCRS.png")
```

@@

```{r ggcont, fig.cap = "A contiguous cartogram map of the United States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each stated shape has been distorted according to the population of the state in 2015. The state of California has become much larger due to it's large population density. This draws attention to the densely populated North East region, and detracts from the Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggcont.png")
```

@@

```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state shape has been maintained, but the size has altered according to the population of the state in 2015. The state of California has remained closer to it's original size than it's surrounding states. The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut. This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggncont.png")
```

```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state has been represented by a circle, but the size was determined by the population of the state in 2015. The North East states remain closer to their neighbours, and may be displaced from their geographic location. The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
knitr::include_graphics("figures/ggdorl.png")
```

@@

He answers this with his implementation of maps created with 'the simplest of all shapes'. While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'. Circular cartograms use the same simple shapes for every region represented, and size them according to the statistic represented or the population for a base map. This familar shape may be more effective that contiguous cartograms, as the shapes used have 'no meaning' after distortions are applied @NISCC.
To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations. This implementation can work for up to 'one hundred thousand' areas.

<!--
-sacrifice communication efficiency;
-There is evidence, too, that angular shapes are more easily seen than ones with smooth outlines. Shapes that contain the qualities of simplicity and regularity are more readily perceived than others that do not possess these qualities.
-Finally, psychologists tell us that shapes having familiarity are more quickly
recognized than nonsense shapes or those that have no meaning.
-The cartographer can efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
-
@NISCC
